---
title: "Intensity Based Genomic prediction models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This Document contains code and data needed to reproduce results from Angarita Barajas et al. Estimation of indirect social genetic effects for skin lesion count in group-housed pigs by quantifying behavioral interactions.

## 1. Load packages
Make sure to put all data in the working folder set below:

Load G matrix (VanRaden,2008), R object: G, class: matrix, Dimension 1079

Phenotypes: Lesion score log(y+1) file, finisher post-mixing, R object: lcount.data
```{r}
rm(list = ls())
#setwd("C:\\Users\\marti\\OneDrive\\Documents\\phenomics\\Belcy\\paper\\data")
setwd("~/Documents/Belcy_MSU_HCPP/2_Files_Matrix_Estimation_Varcomponents_Reml/Files_github_SGE/")
# 1. load library and files and matrix 
library(regress)
library(gwaR)
library(Matrix)
library(MASS)
load("G_matrix.Rdata")
load("lesion_Count_Data.Rdata")
# 1.3.1. Matrix in Reciprocal fight behavior, R object:Zc.RF.Mix, class: matrix
load("Zc_Undirectional_Mix.Rdata")
# 1.3.2. Matrix in Attacks beharior, R object:Zc.ATB.Mix, class: matrix
load("Zc_Directional_Mix.Rdata")
# 1.3.3. Matrix uniform social interaction, R object:Zc.Unif.Big, class: matrix
load("Zc_Uniform_undir_Mix.Rdata")
```
## 2. Estimating (co)variance components with Direct Genetic Effect model (DGE) for one trait (anterior lesions) 

This is the classical animal model implemented by our package gwaR. It includes random effects of pen, fixed effect of sex, rep (batch), pre-mix lesion count, and observers, interaction time in reciprocal fight behavior as covariate and, of course, a genetic additive effect.
```{r}
# 2.1. (co)variance components Anterior lesion count trait
gbfl.covRF<-gblup(rsp = "Finisher_Post_Front_ls", data = lcount.data,
                  design = c(y ~ Sex + Rep + Finisher_Pre_Front_ls + RFight.std +
                               Finisher_Pre_Obs:Finisher_Post_Obs+Finisher_Wt,
                             ~ Finisher_Pen), G=G, pos= c(T,T,T))
varcomp(gbfl.covRF)
```

## 3. Fitting traditional Social Genetic Effect model (TSGE) including interactio time in reciprocal fight as covariate

This are the traditional model with uniform Zc

```{r}
#source("C:\\Users\\marti\\OneDrive\\Documents\\phenomics\\Belcy\\paper\\Functions_Reml_EM.R")
source("Functions_Reml_EM.R")
Zunif.st<-st.Zcmat(Zc.Unif.Big) 

#example of incidence matrix for 2 social groups
heatmap(1/Zc.Unif.Big[1:29,1:29],symm=T,col=gray.colors(32,start=0,end=1),Rowv=NA,Colv = NA,main="Zc assuming uniform interaction")
# fit model Anterior lesion count trait (300 iter)
flcov.tsgeRF<-igest(gbfl.covRF,Zunif.st,tol = 10^-5,k_iter = 300)
# 3.1. Variance-covariance matrix and standard error for estimates 
invImat(flcov.tsgeRF)
# 3.2. Heritability and standard error
varcompreml(flcov.tsgeRF)
# 3.3. Correlation between  direct and social genetic effect and standard error
varcovrml(flcov.tsgeRF)
```
## 4. Fitting intensity-based Social Genetic Effect model (ISGE) for one trait
This are the models proposed in the paper. We first need to load new functions to add the social genetic effects to the base
```{r}
# 4.1 Standardaized interaction social matrix
Zcst.RF<-st.Zcmat(Zc.RF.Mix)  
#example of incidence matrix for 2 social groups, accounting for actual social interactions
heatmap(1/Zc.RF.Mix[1:29,1:29],symm=T,col=gray.colors(32,start=0,end=1),Rowv=NA,Colv = NA,main="Zc for ISGE parameterized with Reciprocal fights")
# fit model (max: 400 iter)
flcov.isgeRF<-igest(gbfl.covRF,Zcst.RF,tol = 10^-5,k_iter = 400)
# 4.1. Variance-covariance matrix and standard error for estimates 
#invImat(flcov.isgeRF)
# 4.2. Heritability and standard error
varcompreml(flcov.isgeRF)
# 4.3. Correlation genetic direct and Indirect and standard error
varcovrml(flcov.isgeRF)
```

